Storm surges in the Strait of Georgia simulated with a regional model

### 2015-01-02 AO_guides/manuscriptAO.tex
edited IB comments

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. Forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
\[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.  The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}. As such, the contribution due to local atmospheric pressure is usually small since the spatial gradient in the sea level pressure is small compared to the model domain. Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
\[ P_s = P_1\left(\gamma\frac{z_1}{T_1} +1 \right)^\frac{g}{\gamma R},\]
where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}.

River input provides a significant volume of fresh water to the Salish Sea and can influence stratification, circulation, and primary productivity. However, most rivers in the domain are not gauged so parametrizations were required to represent river flow. \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation. Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages.

Freshwater runoff from each watershed was divided between the rivers in that watershed. The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website. The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%). The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperature. A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.

>>>>

The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2012 are available. One simulation employs the western Canada component of the High Resolution Deterministic Prediction System (HRDPS), a nested 2.5 km resolution atmospheric model provided by Environment Canada \citep{ECModel}. In addition to momentum and heat fluxes, forcing due to the local atmospheric pressure is also included by calculating an inverse barometer sea surface height,
\[ \eta_{IB} = -\frac{1}{g\rho_{0}}\left(P_{s}-P_0\right),\]
where $g$ is the accelaration due to gravity, $\rho_{0}$ is a reference density set to 1035 kg m$^{-3}$, $P_{s}$ is the sea level atmospheric pressure from the atmospheric model, and $P_0$ is a reference sea level pressure set to 101,000 Pa.  The gradient of $\eta_{IB}$  is scaled by $g$ and added to the forcing terms in the ocean momentum equations \citep{madec2012nemo}. Since the atmospheric model employs a terrain following vertical coordinate the model surface pressure has been corrected to sea level using \citep{holton1992introduction}
\[ P_s = P_1\left(\gamma\frac{z_1}{T_1} +1 \right)^\frac{g}{\gamma R},\]
where $P_1$ and $T_1$ are the model pressure and temperature at height $z_1$, $R$ is the ideal gas constant, and $\gamma$ is an temperature lapse rate set to 0.0098 K m$^{-1}$.

River input provides a significant volume of fresh water to the Salish Sea and can influence stratification, circulation, and primary productivity. However, most rivers in the domain are not gauged so parametrizations were required to represent river flow. \citet{morrison2011rivers} provides a method for estimating freshwater runoff in the Salish Sea region based on precipitation. Monthly runoff volumes for each watershed for each year from 1970 to 2012 were acquired from \citet{morrison2011rivers}, as well as monthly averages. Freshwater runoff from each watershed was divided between the rivers in that watershed. The area drained by each river was estimated from Toporama maps by the Atlas of Canada and watershed maps available on the Washington State government website. The watersheds included in our model were Fraser (which represents approximately 44\% of the freshwater input into our domain), Skagit (12\%), East Vancouver Island (North and South) (12\%), Howe (7\%), Bute (7\%), Puget (6\%), Juan de Fuca (5\%), Jervis (4\%) and Toba (3\%). The monthly flow from each river was input as a point source from 0 to 3 metres at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperature. A total of 150 rivers were parametrized by this method and their locations are indicated by the blue dots in Figure \ref{fig:domain}.
